Measurement Bias from Address Aliasing

### 2014-05-12 paper.tex
Replace data behind microkernel plot and table
Link directly to analysis/../results, with minor post-processing
from parent makefile.

% Large plot of cycle count for two 4K stack sections.
\pgfplotstableread{bin/stack-offset.dat}{\stackoffsettable}
\begin{figure*}[t]
  \caption{Bias from environment size for micro-kernel in Listing~\ref{lst:loopkernel}. Measured average of 100 cycle count samples for 512 different environments. Spikes show aliasing case, occurring once for each 4K period}
  \label{fig:envbias}
  \begin{tikzpicture}
    \begin{axis}[
        width = \textwidth,
        height = 6cm,
        font = \small,
        xlabel=Bytes added to environment,
        ylabel=Cycles,
        domain = 0:8192,
        xtick = {0,1024,...,8192},
        xmin = 0,
        xmax = 8192,
        %legend pos = outer north east,
        cycle list name = exotic
      ]
      %\draw [<->, line width=0.8, color=red] (axis cs:0,8e5) -- (axis cs:4096,8e5) ;
      \addplot[ycomb] table[x expr = \thisrowno{0}*16, y = cycles:u] \stackoffsettable ;
      %\addplot table[x expr = \thisrowno{0}*16, y = r0107:u ] \stackoffsettable ;
      %\addlegendentry{cycles:u} ;
      %\addlegendentry{r0107:u} ;

      % Could annotate area with correlation analysis, (spike - 50*16), + 100*16
      % This is not correct atm, uses 100 samples with byte increment. 
      %\draw [|<->|, gray] (axis cs:2392,720000)--(axis cs:3992,720000) node[right] {Analysis};

    \end{axis}
  \end{tikzpicture}
\end{figure*}

@@

There are clearly two worst cases, indicated by significant spikes at the 3192 and 7288 byte offsets.
We sampled an extensive set of performance counters around the first spike, incrementing environment size by 1 instead of 16 to capture the aliasing case multiple times.
A total of 100 environment offsets were measured, ranging from 3142 to 3242 bytes added.
From this we extracted counters with clear correlation to cycle count.
Table \ref{tab:loopcorrelation} shows numerical counter values for the average case of 3176 bytes added to the environment, compared to worst case of 3192 bytes.
% Events with the highest (absolute value) correlation to cycle count is listed in Table \ref{tab:loopcorrelation}, values close to $1$ or $-1$ indicate strong correlation.

\begin{table}[h]
  \caption{Events with significant correlation to cycle count.}
  \label{tab:loopcorrelation}
  \centering
  \small
  \newcolumntype{H}{>{\lrbox0}c<{\endlrbox}@{}}
  \renewcommand{\tabcolsep}{2pt}
  \begin{tabular}{lHrr}
    \toprule
       &  & 3176 & 3192 \\
    \midrule
      \perfctr{CPU_CLK_UNHALTED.THREAD_P} & cycles:u & 698,101 & 829,103 \\
      \perfctr{CPU_CLK_THREAD_UNHALTED.REF_XCLK} & bus-cycles:u & 19,698 & 23,632 \\ 
      \perfctr{LD_BLOCKS_PARTIAL.ADDRESS_ALIAS} & r0107:u & \bf{121} & \bf{327,908} \\ 
      \perfctr{CYCLE_ACTIVITY.CYCLES_LDM_PENDING} & r02a3:u & 1,183,945 & 1,449,964 \\ 
      \perfctr{RESOURCE_STALLS.ANY} & r01a2:u & 276,648 & 408,071 \\ 
      \perfctr{CPL_CYCLES.RING123} & r025c:u & 690,895 & 832,806 \\ 
      \perfctr{UOPS_EXECUTED_PORT.PORT_2} & r04a1:u & 336,035 & 345,766 \\ 
      \perfctr{CYCLE_ACTIVITY.STALLS_L2_PENDING} & r05a3:u & 492,163 & 512,649 \\ 
      %L2_RQSTS.RFO_MISS & r2224:u & 123 & 127 \\ 
      %OFFCORE_REQUESTS.DEMAND_RFO & r04b0:u & 129 & 130 \\
      %L2_RQSTS.ALL_RFO & re224:u & 245 & 245 \\ 
      %OFFCORE_REQUESTS_OUTSTANDING.DEMAND_RFO & r0460:u & 120154 & 1067959 \\ % Completely random numbers
      %L2_TRANS.RFO & r02f0:u & 249 & 250 \\ 
      %DTLB_STORE_MISSES.STLB_HIT_4K & r2049:u & 294 & 294 \\ 
      %L2_TRANS.L2_WB & r40f0:u & 402 & 395 \\

      %L2_RQSTS.ALL_DEMAND_DATA_RD & re124:u & 1097 & 1091 \\ 
      %MEM_LOAD_UOPS_RETIRED.L2_MISS & r10d1:u & 565 & 562 \\ 
      %L2_RQSTS.RFO_HIT & r4224:u & 69 & 67 \\ 
      %LSD.UOPS & r01a8:u & 1059131 & 1058564 \\

      % Less extreme, but all ports change
      \perfctr{UOPS_EXECUTED_PORT.PORT_3} & r08a1:u & 358,324 & 355,080 \\ 
      \perfctr{UOPS_EXECUTED_PORT.PORT_7} & r80a1:u & 137,862 & 131,882 \\ 
      \perfctr{UOPS_EXECUTED_PORT.PORT_4} & r10a1:u & 280,908 & 276,586 \\

      \perfctr{UOPS_EXECUTED_PORT.PORT_6} & r40a1:u & 234,672 & 180,675 \\ 
      \perfctr{UOPS_EXECUTED_PORT.PORT_0} & r01a1:u & 239,356 & 106,916 \\ 
      \perfctr{UOPS_EXECUTED_PORT.PORT_1} & r02a1:u & 247,572 & 99,270 \\ 
      \perfctr{UOPS_EXECUTED_PORT.PORT_5} & r20a1:u & 250,450 & 122,051 \\ 
      \perfctr{RESOURCE_STALLS.RS} & r04a2:u & 272,833 & 135,900 \\

    \bottomrule

    % \toprule
    %   CPU_CLK_THREAD_UNHALTED.REF_XCLK & 0.99946409726 \\ \midrule
    %   LD_BLOCKS_PARTIAL.ADDRESS_ALIAS & 0.998961259046 \\ \midrule
    %   CYCLE_ACTIVITY.CYCLES_LDM_PENDING & 0.998618840476 \\ \midrule
    %   RESOURCE_STALLS.ANY & 0.998582438059 \\ \midrule
    %   CPL_CYCLES.RING123 & 0.997305209986 \\ \midrule
    %   UOPS_EXECUTED_PORT.PORT_2 & 0.995814514248 \\ \midrule
    %   CYCLE_ACTIVITY.STALLS_L2_PENDING & 0.841261809569 \\ \midrule
    %   UOPS_EXECUTED_PORT.PORT_3 & -0.9780404694 \\ \midrule
    %   UOPS_EXECUTED_PORT.PORT_7 & -0.991034639943 \\ \midrule
    %   UOPS_EXECUTED_PORT.PORT_4 & -0.995703596298 \\ \midrule
    %   UOPS_EXECUTED_PORT.PORT_6 & -0.998681490923 \\ \midrule
    %   UOPS_EXECUTED_PORT.PORT_0 & -0.998757079731 \\ \midrule
    %   UOPS_EXECUTED_PORT.PORT_1 & -0.99886851364 \\ \midrule
    %   UOPS_EXECUTED_PORT.PORT_5 & -0.998956858377 \\ \midrule
    %   RESOURCE_STALLS.RS & -0.998963674679 \\
    % \bottomrule
  \end{tabular}
\end{table}

@@

\begin{sidewaystable*}[p]
  \caption{Most interesting performance counters for O3 convolution code, estimated cost accounting for constant overhead of heap allocation in main. (see /resources for full table)}
  \label{tab:convstats}
  \pgfkeys{/pgf/number format/.cd,int trunc} % Output whole numbers
  \pgfplotstabletypeset[
    font=\normalsize,
    columns={[index]0, [index]1, [index]2, [index]3, [index]4, [index]5, [index]11, [index]21, [index]31}, % Select interesting offsets
    columns/colnames/.style={string type,column name={}, column type={l}},
    columns/colnames/.append style={string replace={cycles:u}{Cycles}}, % Pretty print counter mnemonics
    columns/colnames/.append style={string replace={r0107:u}{Alias}},
    columns/colnames/.append style={string replace={r02a3:u}{cycle_activity.cycles_ldm_pending}},
    columns/colnames/.append style={string replace={r01a2:u}{resource_stalls.any}},
    columns/colnames/.append style={string replace={r05a3:u}{cycle_activity.stalls_l2_pending}},
    columns/colnames/.append style={string replace={r04a2:u}{resource_stalls.rs}},
    columns/colnames/.append style={string replace={r20a1:u}{uops_executed_port.port_5}},
    columns/colnames/.append style={string replace={r0860:u}{offcore_requests_outstanding.all_data_rd}},
    columns/colnames/.append style={string replace={r0160:u}{offcore_requests_outstanding.demand_data_rd}},
    columns/colnames/.append style={string replace={r2049:u}{dtlb_store_misses.stlb_hit_4k}},
    columns/colnames/.append style={string replace={r6049:u}{dtlb_store_misses.stlb_hit}},
    every head row/.style={before row=\toprule, after row=\midrule},
    every last row/.style={after row=\bottomrule}
    ]{\loadedtable}
\end{sidewaystable*}

>>>>

% Large plot of cycle count for two 4K stack sections.
\pgfplotstableread{bin/micro-kernel-cycles.dat}{\stackoffsettable}
\begin{figure*}[t]
  \caption{Bias from environment size for micro-kernel in Listing~\ref{lst:loopkernel}. Measured average of 100 cycle count samples for 512 different environments. Spikes show aliasing case, occurring once for each 4K period}
  \label{fig:envbias}
  \begin{tikzpicture}
    \begin{axis}[
        width = \textwidth,
        height = 6cm,
        font = \small,
        xlabel=Bytes added to environment,
        ylabel=Cycles,
        domain = 0:8192,
        xtick = {0,1024,...,8192},
        xmin = 0,
        xmax = 8192,
        cycle list name = exotic
      ]
      \addplot[ycomb] table[x expr = \thisrowno{0}*16, y = cycles:u] \stackoffsettable ;
    \end{axis}
  \end{tikzpicture}
\end{figure*}

@@

There are clearly two worst cases, indicated by significant spikes at the 3200 and 7296 byte offsets.
We sampled an extensive set of performance counters in addition to cycle count for each execution context.
By looking at the \emph{median} value of each counter compared to the extreme cases, we narrowed down the performance events that behaved similarly to cycle count.
Table \ref{tab:loopcorrelation} shows numerical counter values for median over all environments, compared to worst cases of 3200 and 7296 bytes added.

% Detailed comparison between median and worst cases.
\begin{table}[h]
  \caption{Events with significant correlation to cycle count.}
  \label{tab:loopcorrelation}
  \centering

  \renewcommand{\tabcolsep}{1pt} % Less padding between cells
  \pgfkeys{/pgf/number format/.cd,int trunc} % Output whole numbers

  \pgfplotstabletypeset[
    font=\scriptsize,
    col sep=comma,
    columns={Performance counter, Median, [index]3, [index]4},
    column type=r,
    columns/Performance counter/.style={
      string type, 
      column type=l,
      postproc cell content/.append code={
        \pgfkeysalso{@cell content=\perfctr{##1}}
      }
    },
    every head row/.style={
      output empty row,
      before row={\toprule
        Performance counter & Median & Spike 1 & Spike 2 \\
        },
      after row=\midrule,
    },
    every last row/.style={after row=\bottomrule}
  ]{bin/micro-kernel-comparison.csv}
\end{table}

@@

\begin{sidewaystable*}[p]
  \caption{Most interesting performance counters for O3 convolution code, estimated cost accounting for constant overhead of heap allocation in main.}
  \label{tab:convstats}
  \pgfkeys{/pgf/number format/.cd,int trunc} % Output whole numbers
  \pgfplotstabletypeset[
    font=\normalsize,
    columns={[index]0, [index]1, [index]2, [index]3, [index]4, [index]5, [index]11, [index]21, [index]31}, % Select interesting offsets
    columns/colnames/.style={string type,column name={}, column type={l}},
    columns/colnames/.append style={string replace={cycles:u}{Cycles}}, % Pretty print counter mnemonics
    columns/colnames/.append style={string replace={r0107:u}{Alias}},
    columns/colnames/.append style={string replace={r02a3:u}{cycle_activity.cycles_ldm_pending}},
    columns/colnames/.append style={string replace={r01a2:u}{resource_stalls.any}},
    columns/colnames/.append style={string replace={r05a3:u}{cycle_activity.stalls_l2_pending}},
    columns/colnames/.append style={string replace={r04a2:u}{resource_stalls.rs}},
    columns/colnames/.append style={string replace={r20a1:u}{uops_executed_port.port_5}},
    columns/colnames/.append style={string replace={r0860:u}{offcore_requests_outstanding.all_data_rd}},
    columns/colnames/.append style={string replace={r0160:u}{offcore_requests_outstanding.demand_data_rd}},
    columns/colnames/.append style={string replace={r2049:u}{dtlb_store_misses.stlb_hit_4k}},
    columns/colnames/.append style={string replace={r6049:u}{dtlb_store_misses.stlb_hit}},
    every head row/.style={before row=\toprule, after row=\midrule},
    every last row/.style={after row=\bottomrule}
    ]{\loadedtable}
\end{sidewaystable*}
